---
title: "marine_uk"
output: html_document
date: "2024-07-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

#Install packages
```{r}
install.packages('tidyverse')
library(tidyverse)

install.packages('dplyr')
library(dplyr)

install.packages('naniar')
library(naniar)

install.packages('tidyr')
library(tidyr)

install.packages('stringr')
library(stringr)
```



# Import Tidal and Wave data together, then merge with Energy Prices
```{r}
RUK_tidal_merged <- read_csv("~/Desktop/wave and tidal data/RUK tidal merged.csv")

RUK_wave_merged <- read_csv("~/Desktop/wave and tidal data/RUK wave merged.csv")

```

# Check and clean NA values
```{r}
# TIDAL
is.na(RUK_tidal_merged)

pct_complete(RUK_tidal_merged) #this is from the naniar package   # 5.323383 of this data is complete

tidal_projects <- pull(RUK_tidal_merged, Project)
# drop_na(tidal_projects) did not work #filter(tidal_projects, Project==NA) did not work either
tidal_cleaned <- RUK_tidal_merged %>% 
                 drop_na(Project)
View(tidal_cleaned)

# WAVE

pct_complete(RUK_wave_merged) # 2.00789 of this data is complete

wave_projects <- pull(RUK_wave_merged, Project)

wave_cleaned <- RUK_wave_merged %>% 
                 drop_na(Project)
View(wave_cleaned)

```

# Merge 2 files
```{r}
mergedRUKmarine <- full_join(tidal_cleaned, wave_cleaned)
View(mergedRUKmarine)
```


# Upload with Energy Pricing data (separately)
```{r}
electricity_prices <- read_csv("~/Desktop/wave and tidal data/electricity pricing/electricity_prices.csv")

```

# Clean Energy pricing data to 1) remove NA rows
```{r}
electricity_price_clean <- pull(electricity_prices, year)

electricity_price_clean <- electricity_prices %>% 
                 drop_na(year)
View(electricity_price_clean)
```

# 2) remove NA columns
Note, besides what is listed (dplyr), can also use the janitor package with the following code
data_frame <- data_frame %>%
  remove_empty("cols")
  
```{r}
electricity_price_clean <- electricity_price_clean %>%
  select_if(~ !all(is.na(.)))

View(electricity_price_clean)
```

# 3) Rename [x] values to be actually NA

```{r}

library(naniar)
test <- replace_with_na_all(data = electricity_price_clean,
                    condition = ~.x == "[x]")
View(test)
electricity_price_clean <- test
View(electricity_price_clean)
```

# Merge RUK data with electricity price data

# 1) first extract years specifically from the column is Date_Generator -- this is in dd/mm/yyyy format
# Create a new column for year only
```{r}
# First check the class of the variable
glimpse(mergedRUKmarine)
#Date_Generator is a character column

library(lubridate)

mergedRUKmarine <- mergedRUKmarine %>%
  mutate(
    Date_Generator = dmy(Date_Generator),  # Convert to Date type
    year_only = year(Date_Generator)    # Extract the year; #year_only is the name of the new column
  )
View(mergedRUKmarine)
```

# 2) Merge electricity price data to match the year_only column
```{r}
merged_electric <- mergedRUKmarine %>%
  left_join(electricity_price_clean, by = c("year_only"="year"))
         
View(merged_electric) 
``` 


# Create descriptive stats summary
```{r}
summary(merged_electric)
```
 
# analyze by wave vs. tidal technology (Technology_Generator)
```{r}
tech_grouped <- merged_electric %>% group_by(Technology_Generator)

```
 



# analyze by country

```{r}
merged_electric_grouped <- merged_electric %>% group_by("Country_Generator")

merged_electric_grouped %>% 
  summarize(avg_price= mean(electricty_fuel_price_industry, na.rm = TRUE))
```